---
output:
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
<br>

# **Code Walkthrough**
<br>

### The following page is the step-by-step explanation of the code written in R to create the various charts. 

<br>

#### 1. Load libraries
```{r message=FALSE, warning=FALSE}

library(ggplot2)
library(dplyr)
library(tidyr)
```
<br>

#### 2. Load CSV files and check dataframe structures
```{r echo=TRUE, message=FALSE, warning=FALSE}
list.files(pattern = ".csv")
ward_data <- read.csv("2014_BBMP_Ward_Details.csv")
bus_stops <- read.csv("2012_Bus_Stops.csv")
roads_data <- read.csv("2016_Ward_Roads.csv")

str(ward_data)
str(bus_stops)
str(roads_data)
```
<br>

#### 3. Subset dataframes to separate only required columns
```{r echo=TRUE, message=FALSE, warning=FALSE}
wards <- subset(ward_data, select=c(Ward.No., Ward.Name, Ward.Area..in.sq.km., Total.Population))
colnames(wards) <- c("wardNumber", "wardName", "wardArea", "wardPop")

stops <- subset(bus_stops, select = c(Busstop_Name, Ward_Number))
colnames(stops) <- c("busStops", "wardNos")

roads <- subset(roads_data, select=c(Ward_No, Road.length..kms.))
colnames(roads) <-c("wardNumber", "roadLength")
```
<br>

#### 4. Merge dataframes together to create complete dataframe and clean up final dataframe by removing NA values
```{r echo=TRUE, message=FALSE, warning=FALSE}
roads$wardNumber <- as.numeric(roads$wardNumber)

first_join <- full_join(wards, stops, by=c("wardNumber"="wardNos"))
main_join <- full_join(first_join, roads, by=c("wardNumber"="wardNumber"))

numStops <- main_join %>% count(wardName)

df <- full_join(main_join, numStops, by = "wardName") 

newdf <- subset(df, select=c(wardNumber, wardName, wardArea, wardPop, roadLength,n))
newdf <- na.omit(newdf)
newdf$roadLength <- as.numeric(newdf$roadLength)

str(newdf)
```
<br>

#### 5. Add columns of population density, normalised population, normalised area, and confirm that all rows are unique
```{r echo=TRUE, message=FALSE, warning=FALSE}
finaldf <- newdf %>%  mutate(popDensity = newdf$wardPop/newdf$wardArea) %>% 
  mutate(normPop=(newdf$wardPop-35122)/14213) %>% 
  mutate(normArea=(newdf$wardArea-0.3564)/25.8022)

finaldf <- unique(finaldf)

```
<br>

#### 6. Plot Graphs
```{r cars, echo=TRUE, message=FALSE, warning=FALSE}
## Scatter plot of Ward Area vs. Ward Population

g <- ggplot(data = finaldf, 
            aes(x = wardArea, y = wardPop))
g + geom_point(colour="darkblue", alpha=0.6, size=2) +
  labs(title = "Ward Area vs. Ward Population\n", x = "\nWard Area", y = "Ward Population\n") + 
  theme(plot.margin = margin(15,15,15,15))

```

```{r echo=TRUE, message=FALSE, warning=FALSE}
## Scatter plot of Ward Area vs. Population Density

finaldf$n <- as.numeric(finaldf$n)
h <- ggplot(data = finaldf, 
            aes(x = wardArea, y = popDensity, colour=n))
h + geom_point(size=2, alpha=0.8) + scale_color_gradient(low = "darkblue", high = "orange") + 
  labs(title = "Ward Area vs. Population Density\n", x = "\nWard Area", y = "Population Density\n", colour = "Number of Bus Stops\n(per ward)\n") + 
  theme(plot.margin = margin(15,15,15,15),legend.text = element_text(size = 10), legend.title = element_text(size = 10))

```

```{r echo=TRUE, message=FALSE, warning=FALSE}
### Closer view of subset (<75000 people/sq.km., <10 sq.km.)

j <- ggplot(data = finaldf, 
            aes(x = wardArea, y = popDensity, colour=n))
j + geom_point(size=3, alpha=0.5) + scale_color_gradient(low = "darkblue", high = "orange") + 
  xlim(0,10) + ylim(0, 75000) + 
  labs(title = "Ward Area vs. Population Density\n", x = "\nWard Area", y = "Population Density\n", colour = "Number of Bus Stops\n(per ward)\n")+ 
  theme(plot.margin = margin(15,15,15,15), legend.text = element_text(size = 10), legend.title = element_text(size = 10))

```

```{r echo=TRUE, message=FALSE, warning=FALSE}
## Scatter plot of Population Density vs. Number of Bus Stops (Per ward)

k <- ggplot(data = finaldf, 
            aes(x = popDensity, y = n, color=normArea))
k + geom_point(alpha=1, size=5) + scale_color_gradient(low = "darkblue", high = "orange") +
  geom_point(colour= "black", fill="white", shape=1, size=5, alpha=finaldf$normArea) +
  labs(title = "Population Density vs. Number of Bus Stops (Per ward)\n", x = "\nPopulation Density", y = "Number of Bus Stops\n", colour = "Normalised Ward Area\n")+ 
  theme(plot.margin = margin(15,15,15,15), legend.text = element_text(size = 9), legend.title = element_text(size = 10))
```

```{r echo=TRUE, message=FALSE, warning=FALSE}
### Closer view of subset (<45000 people/sq.km., <13 bus stops)

m <- ggplot(data = finaldf, 
            aes(x = popDensity, y = n, color=normArea))
m + geom_point(alpha=1, size=5) + scale_color_gradient(low = "darkblue", high = "orange") +
  geom_point(colour= "black", fill="white", shape=1, size=5, alpha=finaldf$normArea) + xlim(0,45000) + ylim(0, 13) + 
  labs(title = "Population Density vs. Number of Bus Stops (Per ward)\n", x = "\nPopulation Density", y = "Number of Bus Stops\n", colour = "Normalised Ward Area\n")+ 
  theme(plot.margin = margin(15,15,15,15), legend.text = element_text(size = 9), legend.title = element_text(size = 10))
```

```{r echo=TRUE, message=FALSE, warning=FALSE}
## Scatter plot of Road Lengths vs. Number of Bus Stops (Per ward)

s <- ggplot(data = finaldf, 
            aes(x = roadLength, y = n, color=normArea))
s + geom_point(size=4, alpha=0.8) + scale_color_gradient(low = "darkblue", high = "orange")+
  labs(title = "Road Lengths vs. Number of Bus Stops (Per ward)\n", x = "\nRoad Lengths", y = "Number of Bus Stops\n", colour = "Normalised Ward Area\n")+ 
  theme(plot.margin = margin(15,15,15,15), legend.text = element_text(size = 9), legend.title = element_text(size = 10))
```

```{r echo=TRUE, message=FALSE, warning=FALSE}
### Closer view of subset (<100 km., <12 bus stops)

t <- ggplot(data = finaldf, 
            aes(x = roadLength, y = n, color=normArea))
t + geom_point(size=4, alpha=0.8) + scale_color_gradient(low = "darkblue", high = "orange") + xlim(0,100) + ylim(0,12) +
  labs(title = "Road Lengths vs. Number of Bus Stops (Per ward)\n", x = "\nRoad Lengths", y = "Number of Bus Stops\n", colour = "Normalised Ward Area\n")+ 
  theme(plot.margin = margin(15,15,15,15), legend.text = element_text(size = 9), legend.title = element_text(size = 10))

```

